---
title: "CRAZIE Data Conversions"
author: "Zechariah Meunier"
date: "Last updated `r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---
### Libraries and Data

Load libraries:  

* `dplyr` for data manipulations  
* `reshape2` for restructuring data frames  
* `vegan` for dissimilarity indices and Mantel tests
```{r message=FALSE}
library(dplyr)
library(reshape2)
library(vegan)
```

Load plot descriptor matrix (`dscmat`), species matrix (`sppmat`), and species to functional group mapping matrix (`spp2FGmat`)
```{r}
dscmat <- read.csv(file="../Data/CRAZIE Descriptor Matrix.csv",
                   header=TRUE, sep=",")

sppmat <- read.csv(file="../Data/CRAZIE Species Matrix.csv",
                   header=TRUE, sep=",", row.names = 1)

spp2FGmat <- read.csv(file="../Data/CRAZIE Spp-FG Mapping Matrix.csv",
                      header=TRUE, sep=",")
```

Subset surveyed plots and years of interest from `dscmat` and `sppmat`.
```{r}
years <- c(2006:2020)

dscmat_subset <- filter(dscmat, Surveyed == 1, Year %in% years)
dscmat_unsurveyed <- filter(dscmat, Surveyed == 0, Year %in% years)

sppmat_subset <- filter(sppmat, row.names(sppmat) %in% dscmat_subset$PlotID)
```

Create functional group matrices by organizing taxa into specified functional groups and summing counts and percent covers within groups.
```{r}
#Narrow Functional Group Matrix
narrowFGmat <- sppmat_subset

colnames(narrowFGmat) <- spp2FGmat$NarrowFG2[match(colnames(sppmat_subset),
                                            spp2FGmat$ColName)]

narrowFGmat <- as.data.frame(t(rowsum(t(narrowFGmat), group = colnames(narrowFGmat),
                                reorder = F, na.rm = T)))

#Broad Functional Group Matrix
broadFGmat <- sppmat

colnames(broadFGmat) <- spp2FGmat$BroadFG[match(colnames(sppmat_subset),
                                            spp2FGmat$ColName)]

broadFGmat <- as.data.frame(t(rowsum(t(broadFGmat), group = colnames(broadFGmat),
                                reorder = F, na.rm = F)))
```

Subset functional groups of interest.
```{r}
broadFGs <- c("Substrate", "Sessile invertebrates", "Crusts", "Macrophytes")

narrowFGs <- unique(subset(spp2FGmat, BroadFG %in% broadFGs)$NarrowFG2)

sslFGmat <- select(narrowFGmat, all_of(narrowFGs))

trtFGs <- c("Sessile invertebrates", "Macrophytes")

trtFGmat <- select(broadFGmat, all_of(trtFGs))
```

### Compare Community Matrices 

Calculate coefficient of variation, relativizations, and transformations.
```{r}
#Coefficient of variation for columns and rows. Values >100 suggest that 
#relativizing could have a large effect on results.
paste("Column CV%:", round(100*sd(colSums(sslFGmat))/mean(colSums(sslFGmat)),
                           digits = 2))
paste("Row CV%:", round(100*sd(colSums(t(sslFGmat)))/mean(colSums(t(sslFGmat))),
                           digits = 2))

#Relativization by column maximum for proportion data
source("Matrix Relativizing Functions.R")
sslFG_prop <- rel.col.max(sslFGmat)

#Transformations for percent cover data
sslFG_sqrt <- sqrt(sslFGmat)
sslFG_4rt <- sslFGmat^0.25
sslFG_log <- log(sslFGmat+1)

#Transformations for proportion data
sslFG_prop_sqrt <- sqrt(sslFG_prop)
sslFG_prop_asin_sqrt <-2/pi*asin(sqrt(sslFG_prop))
```

Calculate dissimilarity indices for all matrices.
```{r}
#Bray-Curtis dissimilarities
sslFG_BC <- as.matrix(vegdist(sslFGmat, method = "bray"))
sslFG_sqrt_BC <- as.matrix(vegdist(sslFG_sqrt, method = "bray"))
sslFG_4rt_BC <- as.matrix(vegdist(sslFG_4rt, method = "bray"))
sslFG_log_BC <- as.matrix(vegdist(sslFG_log, method = "bray"))
sslFG_prop_BC <- as.matrix(vegdist(sslFG_prop, method = "bray"))
sslFG_prop_sqrt_BC <- as.matrix(vegdist(sslFG_prop_sqrt, method = "bray"))
sslFG_prop_asin_sqrt_BC <- as.matrix(vegdist(sslFG_prop_asin_sqrt, method = "bray"))

#Chao dissimilarities
sslFG_Chao <- as.matrix(vegdist(round(sslFGmat, digits = 0), method = "chao"))
```

Perform Mantel tests to compare matrices. If tests are significant, then reject the null hypothesis that there is no relationship between the two matrices.
```{r}
mantel(sslFG_BC, sslFG_Chao)
mantel(sslFG_BC, sslFG_sqrt_BC)
mantel(sslFG_BC, sslFG_4rt_BC)
mantel(sslFG_BC, sslFG_log_BC)
mantel(sslFG_BC, sslFG_prop_BC)
mantel(sslFG_BC, sslFG_prop_sqrt_BC)
mantel(sslFG_BC, sslFG_prop_asin_sqrt_BC)
```

Reformat Bray-Curtis dissimilarity matrix from square to pairwise.
```{r}
idx <- lower.tri(sslFG_BC)
sslFG_BC_long <- cbind(rev(expand.grid(dimnames(sslFG_BC)))[which(idx),], val = sslFG_BC[idx])
colnames(sslFG_BC_long) <- c("PlotID1","PlotID2","BC.Dist")
```

