---
title: "CRAZIE Marine Heatwave"
author: "Zechariah Meunier"
date: "Last updated `r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

Load libraries:

* `ggplot2` for creating data visualizations
* `dplyr` for data frame manipulations
* `heatwaveR` for detecting marine heatwaves
* `lubridate` for working with dates
* `extrafont` for custom fonts
```{r, include=FALSE}
library(ggplot2)
library(dplyr)
library(heatwaveR)
library(lubridate)
library(extrafont)
```

Store daily mean sea surface temperature (SST) per site and calculate daily mean SST per cape.
```{r}
SST_site <- CRZdailySST
SST_cape <- vector(mode = "list", length = 4)
names(SST_cape) <- c("CF","CP","CB","CM")

SST_cape[[1]] <- filter(dailySSTlong, site %in% c("FC","BB","MB")) %>%
  group_by(t) %>%
  summarize(temp = mean(temp, na.rm = TRUE))

SST_cape[[2]] <- filter(dailySSTlong, site %in% c("YB","SH","TK")) %>%
  group_by(t) %>%
  summarize(temp = mean(temp, na.rm = TRUE))

SST_cape[[3]] <- filter(dailySSTlong, site %in% c("CBN","CBS","POH","RP")) %>%
  group_by(t) %>%
  summarize(temp = mean(temp, na.rm = TRUE))

SST_cape[[4]] <- filter(dailySSTlong, site %in% c("CMN","CMS","KH")) %>%
  group_by(t) %>%
  summarize(temp = mean(temp, na.rm = TRUE))
```

Calculate climatologies and thresholds (`clim`) from temperature series. 
Set `pctile = 90` for marine heatwaves or `10` for coldspells.
```{r}
#site-level
clim_site <- list()

for(i in 1:13){
  clim_site[[i]] <- ts2clm(SST_site[[i]], x=t, y=temp, pctile=90,
               climatologyPeriod = c(min(SST_site[[i]]$t),max(SST_site[[i]]$t)))
  names(clim_site)[i] <- names(SST_site)[i]}

#cape-level
clim_cape <- list()

for(i in 1:4){
  clim_cape[[i]] <- ts2clm(SST_cape[[i]], x=t, y=temp, pctile=90,
               climatologyPeriod = c(min(SST_cape[[i]]$t),max(SST_cape[[i]]$t)))
  names(clim_cape)[i] <- names(SST_cape)[i]
  write.csv(clim_cape[[i]], paste0("../Results/R Export/",
                                    names(SST_cape[i]), " Climatology.csv"))}

#only run below if loading climatologies already calculated
clim_cape[[1]] <- read.csv("../Results/R Export/CF Climatology.csv")[,2:6]
clim_cape[[2]] <- read.csv("../Results/R Export/CP Climatology.csv")[,2:6]
clim_cape[[3]] <- read.csv("../Results/R Export/CB Climatology.csv")[,2:6]
clim_cape[[4]] <- read.csv("../Results/R Export/CM Climatology.csv")[,2:6]
names(clim_cape) <- c("CF","CP","CB","CM")
clim_cape[[1]]$t <- date(clim_cape[[1]]$t)
clim_cape[[2]]$t <- date(clim_cape[[2]]$t)
clim_cape[[3]]$t <- date(clim_cape[[3]]$t)
clim_cape[[4]]$t <- date(clim_cape[[4]]$t)
```

Detect extreme temperature events (`events`) and save graph (`x_graph`). 
Set `coldSpells = FALSE` for heatwaves or `TRUE` for coldspells.
```{r}
for(i in 1:13){
  site = names(clim_site)[i]
  events <- detect_event(clim_site[[site]], coldSpells = FALSE) 

  all_graph <- event_line(events, spread = 5000, category = FALSE)
  extreme_graph <- event_line(events, spread = 600, category = FALSE)
  category_graph <- event_line(events, spread = 250, category = TRUE)
  blob_graph <- event_line(events, start_date = "2014-01-01",
                           end_date = "2016-08-31", category = TRUE)

  ggsave(all_graph, width = 6, height = 4, units = "in", dpi = 2000,
       filename = paste0("../Figures & Tables/MHW Graphs/",site,"_all.jpg"))
  ggsave(extreme_graph, width = 6, height = 4, units = "in", dpi = 2000,
       filename = paste0("../Figures & Tables/MHW Graphs/",site,"_extreme.jpg"))
  ggsave(category_graph, width = 6, height = 4, units = "in", dpi = 2000,
       filename = paste0("../Figures & Tables/MHW Graphs/",site,"_category.jpg"))
  ggsave(blob_graph, width = 6, height = 4, units = "in", dpi = 2000,
       filename = paste0("../Figures & Tables/MHW Graphs/",site,"_blob.jpg"))}
```

Save theme for marine heatwave (MHW) graph.
```{r}
blobtheme <- 
  theme(axis.title = element_text(size=10, color = "gray10", family = "Roboto"),
        axis.text.y = element_text(color = "gray10", family = "Roboto",
                                   margin=margin(l=-2, r=4, unit = "pt")),
        axis.text.x = element_text(color = "gray10", family = "Roboto",
                                   margin=margin(t=4, unit = "pt")),
        axis.ticks = element_line(color = "gray10"),
        axis.ticks.length = unit(3, "pt"),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA, color = "gray10", size = 1),
        legend.key = element_blank(), 
        legend.position = "none",
        legend.box.spacing = unit(0.1,"cm"),
        legend.title = element_text(size = 10, color = "gray10",
                                    family = "Roboto Medium"),
        legend.text = element_text(size = 9, color = "gray10",
                                   family = "Roboto"),
        plot.tag = element_text(size = 13, color="gray10", family = "Roboto"),
        plot.tag.position = c(0.01,0.99))

capetags <- c("a","b","c","d")
capelabs <- c("Cape Foulweather, OR",
              "Cape Perpetua, OR      ",
              "Cape Blanco, OR          ",
              "Cape Mendocino, CA  ")
```

Create MHW graphs per cape.
```{r}
for(i in 1:4){
  cape = names(clim_cape)[i]
  events <- detect_event(clim_cape[[cape]], coldSpells = FALSE) 

  all_graph <- event_line(events, spread = 5000, category = FALSE)
  extreme_graph <- event_line(events, spread = 600, category = FALSE)
  blob_graph <- event_line(events, spread = 1000, category = FALSE,
                           y_axis_title = "Temperature (\u00B0C)") +
    annotate("text", color = "gray10", size = 2.9,
           x = as.Date("2014-03-22", "%Y-%m-%d"), y = 17,
           label = capelabs[i],
           family = "Roboto") +
   # annotate("segment", x = as.Date("2014-08-01", "%Y-%m-%d"),
    #         xend = as.Date("2015-05-31", "%Y-%m-%d"), y = 7.5, yend = 7.5,
    #       arrow = arrow(ends = "both", angle = 90, length = unit(.2,"cm"))) +
    annotate("rect", alpha = 0, color = "gray10", linetype = "dashed", 
             xmin = as.Date("2014-08-01", "%Y-%m-%d"), 
             xmax = as.Date("2015-05-31", "%Y-%m-%d"), 
             ymin = 7.5, ymax = 17.5) +
    coord_fixed(ratio = 30, expand = FALSE, ylim = c(5.95,18.05),
                xlim = c(as.Date("2013-11-01", "%Y-%m-%d"),
                         as.Date("2017-02-28", "%Y-%m-%d"))) +
    blobtheme #+ labs(tag = capetags[i])
    

  #ggsave(all_graph, width = 6, height = 4, units = "in", dpi = 2000,
   #    filename = paste0("../Figures & Tables/MHW Graphs/",cape,"_all.jpg"))
  #ggsave(extreme_graph, width = 6, height = 4, units = "in", dpi = 2000,
   #    filename = paste0("../Figures & Tables/MHW Graphs/",cape,"_extreme.jpg"))
  ggsave(blob_graph, width = 6, height = 2, units = "in", dpi = 2000,
       filename = paste0("../Figures & Tables/MHW Graphs/",cape,"_blob20240330.svg"))}
```

Make the category MHW graphs from scratch.
```{r}
#Set line colors
lineColCat <- c(
  "Temperature" = "black",
  "Climatology" = "blue",
  "Threshold" = "darkgreen",
  "2x Threshold" = "darkgreen",
  "3x Threshold" = "darkgreen",
  "4x Threshold" = "darkgreen"
  )

#Set category fill colors
fillColCat <- c(
  "Moderate" = "#ffc866",
  "Strong" = "#ff6900",
  "Severe" = "#9e0000",
  "Extreme" = "#2d0000"
  )

for(i in 1:4){
  cape = names(clim_cape)[i]
  events <- detect_event(clim_cape[[cape]], coldSpells = FALSE) 

#Create category breaks
clim_cat <- events$climatology %>%
  dplyr::mutate(diff = thresh - seas,
                thresh_2x = thresh + diff,
                thresh_3x = thresh_2x + diff,
                thresh_4x = thresh_3x + diff)

#Create category plots
category_graph <- ggplot(data = clim_cat, aes(x = t, y = temp)) +
  geom_flame(aes(y2 = thresh, fill = "Moderate")) +
  geom_flame(aes(y2 = thresh_2x, fill = "Strong")) +
  geom_flame(aes(y2 = thresh_3x, fill = "Severe")) +
  geom_flame(aes(y2 = thresh_4x, fill = "Extreme")) +
  geom_line(aes(y = thresh_2x, col = "2x Threshold"), size = 0.7, linetype = "dashed") +
  geom_line(aes(y = thresh_3x, col = "3x Threshold"), size = 0.7, linetype = "dotdash") +
  geom_line(aes(y = thresh_4x, col = "4x Threshold"), size = 0.7, linetype = "dotted") +
  geom_line(aes(y = seas, col = "Climatology"), size = 0.7) +
  geom_line(aes(y = thresh, col = "Threshold"), size = 0.7) +
  geom_line(aes(y = temp, col = "Temperature"), size = 0.6) +
  scale_colour_manual(name = NULL, values = lineColCat,
                      breaks = c("Temperature", "Climatology", "Threshold",
                                 "2x Threshold", "3x Threshold", "4x Threshold")) +
  scale_fill_manual(name = NULL, values = fillColCat, guide = FALSE) +
  scale_x_date(date_labels = "%b %Y") +
  guides(colour = guide_legend(override.aes = list(linetype = c("solid", "solid", "solid",
                                                                "dashed", "dotdash", "dotted"),
                                                   size = c(0.6, 0.7, 0.7, 0.7, 0.7, 0.7)))) +
  labs(y = NULL, x = NULL) +
  coord_fixed(ratio = 15, expand = FALSE, ylim = c(7.5,21.5),
                xlim = c(as.Date("2014-08-01", "%Y-%m-%d"),
                         as.Date("2015-05-31", "%Y-%m-%d"))) + 
  blobtheme + theme(legend.position = "none")

ggsave(category_graph, width = 3, height = 2.5, units = "in", dpi = 2000,
       filename = paste0("../Figures & Tables/MHW Graphs/",cape,"_category3.svg"))
}
```


